Missing Data Patterns

### 2025-03-06 notebooks/analysis.qmd
Initialising gh-pages branch

```{r}
#| label: tbl-avail
#| tbl-cap: Data availability by year.
hsp |>
  mutate(
    year = year(date),
    missing_price = is.na(price),
    missing_type = is.na(type),
    missing_tenure = is.na(tenure),
    missing_status = is.na(status),
    missing_plot_area = is.na(plot_area),
    missing_floor_area = is.na(floor_area),
    missing_storeys = is.na(storeys),
    missing_beds = is.na(beds),
    missing_baths = is.na(baths),
    missing_housechar = missing_plot_area & missing_floor_area & missing_beds & missing_baths,
    year = year(quarter),
  ) |>
  summarise(
    count = n(),
    spatial = length(unique(mukim)) / length(unique(hsp$mukim)),
    # price = sum(missing_price),
    type = sum(missing_type),
    house_char = sum(missing_housechar),
    source = list(unique(method)),
    .by = year
  ) |>
  mutate(
    `nat-archive` = map_chr(source, ~ ifelse("nat-archive" %in% .x, "✔", "")),
    `online-archive` = map_chr(source, ~ ifelse("online-archive" %in% .x, "✔", "")),
    `web-scrape` = map_chr(source, ~ ifelse("web-scrape" %in% .x | "web-scrape-llm" %in% .x, "✔", "")),
    llm = map_chr(source, ~ ifelse("web-scrape-llm" %in% .x, "✔", ""))
  ) |> 
  select(-source) |>
  mutate(across(type:house_char, \(x) x / count)) |>
  gt(rowname_col = "year") |>
  cols_align(
    align = "center",
    columns = `nat-archive`:llm
  ) |>
  tab_spanner(
    label = "Data source",
    columns = `nat-archive`:`web-scrape`
  ) |>
  tab_spanner(
    label = "Missing data severity",
    columns = type:house_char
  ) |>
  fmt_percent(
    spatial:house_char,
    decimals = 1
  ) |>
  cols_label(
    year ~ "Year",
    count ~ "Count",
    spatial ~ "Spatial coverage (mukim)",
    type ~ "Property Type",
    house_char ~ "Property Characteristics",
    `nat-archive` ~ "National Archive",
    `online-archive` ~ "Online Archive",
    `web-scrape` ~ "Web Scraping",
    llm ~ "LLM post-processing"
  ) |>
  grand_summary_rows(
    columns = count:house_char,
    fns = list(Mean = "mean"),
    fmt = list(
      ~ fmt_number(., decimals = 0, columns = "count"),
      ~ fmt_percent(., decimals = 1, columns = c("spatial", "type", "house_char"))
    )
  ) |>
  tab_footnote(
    footnote = paste0("Of Brunei’s 39 mukims, only 27 are considered transactable—excluding water villages and remote, non-developable areas."),
    locations = cells_column_labels(columns = spatial)
  ) |>
  tab_footnote(
    footnote = "Unknown property type.",
    locations = cells_column_labels(columns = type)
  ) |>
  tab_footnote(
    footnote = "Missing all of plot area, floor area, beds, and baths variables.",
    locations = cells_column_labels(columns = house_char)
  ) |>
  tab_options(
    quarto.disable_processing = TRUE,
    grand_summary_row.text_transform = "capitalize"
  ) 
```

>>>>

```{r}
#| label: tbl-avail
#| tbl-cap: Data availability by year.
hsp |>
  mutate(
    year = year(date),
    missing_price = is.na(price),
    missing_type = is.na(type),
    missing_tenure = is.na(tenure),
    missing_status = is.na(status),
    missing_plot_area = is.na(plot_area),
    missing_floor_area = is.na(floor_area),
    missing_storeys = is.na(storeys),
    missing_beds = is.na(beds),
    missing_baths = is.na(baths),
    missing_housechar = missing_plot_area & missing_floor_area & missing_beds & missing_baths,
    year = year(quarter),
  ) |>
  summarise(
    count = n(),
    spatial = length(unique(mukim)) / length(unique(hsp$mukim)),
    # price = sum(missing_price),
    type = sum(missing_type),
    house_char = sum(missing_housechar),
    source = list(unique(method)),
    .by = year
  ) |>
  mutate(
    `nat-archive` = map_chr(source, ~ ifelse("nat-archive" %in% .x, "✔", "")),
    `online-archive` = map_chr(source, ~ ifelse("online-archive" %in% .x, "✔", "")),
    `web-scrape` = map_chr(source, ~ ifelse("web-scrape" %in% .x | "web-scrape-llm" %in% .x, "✔", "")),
    llm = map_chr(source, ~ ifelse("web-scrape-llm" %in% .x, "✔", ""))
  ) |> 
  select(-source) |>
  mutate(across(type:house_char, \(x) x / count)) |>
  gt(rowname_col = "year") |>
  cols_align(
    align = "center",
    columns = `nat-archive`:llm
  ) |>
  tab_spanner(
    label = "Data source",
    columns = `nat-archive`:`web-scrape`
  ) |>
  tab_spanner(
    label = "Missing data severity",
    columns = type:house_char
  ) |>
  fmt_percent(
    spatial:house_char,
    decimals = 1
  ) |>
  cols_label(
    year ~ "Year",
    count ~ "Count",
    spatial ~ "Spatial coverage (mukim)",
    type ~ "Property Type",
    house_char ~ "Property Characteristics",
    `nat-archive` ~ "National Archive",
    `online-archive` ~ "Online Archive",
    `web-scrape` ~ "Web Scraping",
    llm ~ "LLM post-processing"
  ) |>
  grand_summary_rows(
    columns = count:house_char,
    fns = list(Mean = "mean"),
    fmt = list(
      ~ fmt_number(., decimals = 0, columns = "count"),
      ~ fmt_percent(., decimals = 1, columns = c("spatial", "type", "house_char"))
    )
  ) |>
  tab_footnote(
    footnote = paste0("Of Brunei’s 39 mukims, only 27 are considered transactable--excluding water villages and remote, non-developable areas."),
    locations = cells_column_labels(columns = spatial)
  ) |>
  tab_footnote(
    footnote = "Unknown property type.",
    locations = cells_column_labels(columns = type)
  ) |>
  tab_footnote(
    footnote = "Missing all of plot area, floor area, beds, and baths variables.",
    locations = cells_column_labels(columns = house_char)
  ) |>
  tab_options(
    quarto.disable_processing = TRUE,
    grand_summary_row.text_transform = "capitalize"
  ) 
```

### 2025-03-06 notebooks/missing.qmd
Initialising gh-pages branch

>>>>

## Main article

```{r}
#| message: false
#| warning: false
#| echo: false
#| label: fig-missing-pattern
#| fig-cap: Missing data patterns for key house characteristics.
#| out-width: 100%
#| fig-width: 9
#| fig-height: 4
hsp |>
  dplyr::select(price, plot_area, floor_area, beds, baths) |>
  plot_pattern(npat = 16) +
  coord_flip() +
  theme(aspect.ratio = 0.3)
```
